Elimination Of Jump Objects

### 2022-10-10 paper/main.tex
Create main.tex

\documentclass[two column]{article}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{graphicx}

\graphicspath{ {./images/} }

% Keywords command
\providecommand{\keywords}[1]
{
  \small
  \textbf{\textit{Keywords:}} #1
}

\title{\textbf{Elimination Of Jump Statements}}
\author{\textbf{author}}
\date{May 2022}

@@

\begin{abstract}
In this article we raise the problem of the so-called Jump Statements. Getting rid of the use of operators such as \textbf{goto}, \textbf{return}, \textbf{continue}, \textbf{break}, and \textbf{throw} in code is the task of most modern programming languages, as the program becomes unreadable, poorly optimized and the probability of fixing errors in such a program is extremely high. 
We are trying to answer the following research question: “Is it possible to completely get rid of the \textbf{GOTO} operator in the code?”

\end{abstract}

\keywords{Object-Oriented Programming, software engineering, algorithms, computer science}

\section{Introduction}
The problem is known both computer science and software engineering communities for a few decades already, while it was firstly formulated by Dijkstra in 1968: he affirmed that the use of \textbf{goto} has disastrous effect and proposed to remove it from all programming languages, Even though a few years later Knuth showed a manner for structuring programs with the \textbf{goto} statement and argued that it could be a powerful tool if it is well used, researchers got interested in finding a way for restructuring programs and eliminating \textbf{goto} statements.

Williams created a tool for restructuring Pascal code; Jonsson was one of the first who made a summary of so called \textbf{GOTO}-patches, which are design techniques for eliminating \textbf{goto} statements; Erosa suggested an algorithm fir automatic elimination of \textbf{goto} statements in C; Zegour explained how \textbf{goto} elimination may work in any language, as long as it has jump statements and control structures; Morris suggested to use regular expressions for replacing \textbf{goto} with IF-THEN-ELSE constructs; Ramshow proposed an algorithm for Pascal programs; Ganapathi introduced optimisation method for SHIM IR; Ceccato suggested how to eliminate \textbf{goto} during migration of legacy code to Java,

To our knowledge, there is still no software tool available, which would implement any of the suggested methods and enable automatic elimination of jump statements in programs. However, the main reason for finding a solution to this problem for us is the development of EOLANG, an experimental \(\varphi\)-calculus-based object-oriented programming language, which is used as an Intermediate Representation (IR) in Polystat, a new polyglot static analyzer. The method of program analysis used in Polystat is functional: we find bugs by 1) transforming source code to formulas of \(\varphi\)-calculus, 2) mapping them to \(\lambda\)-terms, and 3) reasoning on terms using their inverses.

Such an analysis works well with statements which are easy to map to functions, for example if and switch. Jump statements, on the other hand, can't be easily mapped to functions, The presence of jump statements in the source code makes it impossible to use the functional method of analysis just explained.

The result of our research is an algorithm that allows converting a program in EO (Elementary Objects) language containing \textbf{GOTO} objects into a semantically similar one, but without using them. We propose an algorithm that replaces code fragments using the GOTO object with similar fragments, but without using it. We claim that two algorithms are semantically equivalent if with the same set of input arguments they return the same results.

\section{Methods}
In EOLANG, all jump statements are implemented as a single \textbf{GOTO}-object. The \textbf{GOTO}-object represents two types of jumps: \textit{forward jump} and \textit{backward jump}. Operators such as \textbf{break}, \textbf{return} and \textbf{goto forward} are implemented via \textit{forward jump}; \textbf{continue} and \textbf{goto backward} are implemented via \textit{backward jump}.

For each of them, a semantically identical replacement is required so that the algorithm returns the same output with the same input parameters.

\subsection{forward jump}
\textbf{\textit{Goto forward}} is implemented in the language as:
\begin{lstlisting}
+import org.eolang.gray.goto
+import org.eolang.io.stdout

goto
 [g]
  seq
   g.forward TRUE
   stdout "Will never be printed"
\end{lstlisting}

In a program \textit{forward jump} can be used within the body of \textbf{if}-statement. In this case program will look like this:

\includegraphics{forward1}

Let's look at the following code on EOLANG:
\begin{lstlisting}
[x] > f
 memory > r
 seq > @
  r.write 0
  goto
   [g]
    seq > @
     if.
      x.eq 0
      g.forward
      TRUE
     r.write (42.div x)
  r
\end{lstlisting}

In this example we can see that the use of "forward jump" is unnecessary because it is called when condition within \textbf{if}-statement is TRUE. By inverting the condition in \textbf{if}-statement, you can write the code block between the \textbf{goto} call and its label into a separate \textbf{if}-statement:

\includegraphics{forward2}

Modified code:
\begin{lstlisting}
[x] > f
 memory > r
 seq > @
  r.write 0
  if.
   not. (eq. (x 0))
   r.write (42.div x)
   TRUE
  r
\end{lstlisting}

If the program encountered unconditional \textit{forward jump}, we need to transform it as \textbf{goto} with an additional condition: \textbf{if (true) goto forward}. Then the replacement occurs just as if the condition exists.

\subsubsection{return}
As mentioned above, return statement is also implemented via \textit{goto forward} object. Consider the following code:
\begin{lstlisting}
[a b] > max
 goto > @
  [g]
   seq > @
    if.
     a.greater b
     g.forward a
     TRUE
    b
\end{lstlisting}
When using goto as \textit{return} statement, an additional attribute is added to the \textbf{goto}-object, similar to the return value of a procedure. In this case, it can easily be transformed into an \textbf{if}-statement as shown in the paragraph above:
\begin{lstlisting}
[a b] > max
 if.
  a.greater b
  a
  TRUE
 b
\end{lstlisting}

\subsubsection{break}
Break is a control-flow statement which could also be implemented using \textbf{goto forward}. Since the break statement can only be found inside the generator (in EOlang Generator is a routine that can be used to control iteration behaviour of a loop) Its replacement is slightly different from the one presented in the paragraph above. In this case, we need to invert the \textit{if}-condition and add it to the condition of the loop with the logical operator \textit{AND}. After that, we easily get rid of the \textit{if}-condition itself.

Let's look at the example:
\begin{lstlisting}
[]
 memory > i
 goto > @
  [g]
   seq > @
    while.
     i.less 10
     seq
      i.write (i.add 1)
      if.
       eq. (i.mod 5 0)
       g.forward
       TRUE
\end{lstlisting}

We invert \textit{condition} and add it to \textbf{while}-condition:
\begin{lstlisting}
[]
 memory > i
 seq > @
  while.
   and.
    i.less 10
    not. (eq. (i.mod 5 0))
   seq
    i.write (i.add 1)
\end{lstlisting}

\subsection{backward jumps}
Backward jump is usually used similarly to a while loop and can be easily replaced by it (considering programs that are compiled and executed). Substitution will look like \textbf{do-while} construction.

\includegraphics{backward1}

Consider the following code:
\begin{lstlisting}
[] > f
 memory > i
 seq > @
  i.write 1
  goto
   [g]
    seq > @
     i.write (i.add 1)
     if.
      i.less 10
      g.backward
      TRUE
 stdout "Finished!"
\end{lstlisting}

In this case we could easily replace construction using \textbf{goto} with the \textbf{while}:

\includegraphics{backward2}

\begin{lstlisting}
[] > f
 memory > i
 seq > @
  i.write 1
  i.write (i.add 1)
  while.
   i.less 10
   seq
    i.write (i.add 1)
 stdout "Finished!"
\end{lstlisting}

\subsubsection{continue}
Continue is also a control-flow statement, which occurs in loops and is implemented through the EO object \textbf{GOTO}. In order to replace it, we need to invert the \textit{if}-condition in the loop and all the statements that occur after \textbf{goto} are put in the \textit{if} body with the inverted condition.

Example of using \textit{continue} via \textbf{goto backward}:
\begin{lstlisting}
[]
 memory > i
 goto > @
  [g]
   seq > @
    while.
     i.less 10
     seq
      i.write (i.add 1)
      if.
       eq. (i.mod 5 0)
       g.backward
       TRUE
      stdout "Not divisible by 5!\n"
\end{lstlisting}

Replace the GOTO object with the described method:
\begin{lstlisting}
[]
 memory > i
 seq > @
 while.
  i.less 10
  seq
   i.write (i.add 1)
   if.
    not. (eq. (i.mod 5 0))
    stdout "Not divisible by 5!\n"
    TRUE
\end{lstlisting}

\subsection{Complex cases}
??? when forward and then backward jump
??? and mb add throw ???

\subsection{Final algorithm}
Knowing how to replace a goto-forward statement and a goto-backward statement, we can present a final algorithm to get rid of all the GOTO objects in the EOlang code.

\end{document}
\endinput

>>>>

\documentclass[sigplan,review,11pt,nonacm,natbib=false]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\usepackage[maxnames=1,minnames=1,maxbibnames=100,natbib=true,citestyle=authoryear,bibstyle=authoryear,doi=false,url=false,isbn=false,isbn=false,backend=biber]{biblatex}
\usepackage{debate}
\usepackage[T2A,T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel}
  \renewcommand\ttdefault{cmtt}
\usepackage{csquotes}
\usepackage{ffcode}
\usepackage{CJKutf8}
\usepackage{paralist}
\usepackage{cancel} % to enable \cancel command
\usepackage{anyfontsize} % To get rid of font not found warnings
\usepackage{tabularx} % for special tables
\usepackage{amsmath}
\usepackage{to-be-determined}
\usepackage{href-ul}
\usepackage{amsthm}
\usepackage{cleveref}
\usepackage{tcolorbox} % for algorithm
  \tcbuselibrary{skins}
\usepackage{fdsymbol} % for \mathbb
\usepackage{algpseudocode} % for algorithms
\usepackage{multicol} % for two cols in BNF
\usepackage{pgffor} % to enable \foreach
\usepackage{stmaryrd}
\usepackage{mathtools}
\newtheoremstyle{theorems}
  {} % Space above
  {} % Space below
  {} % Theorem body font % (default is "\upshape")
  {} % Indent amount
  {\scshape} % Theorem head font % (default is \mdseries)
  {.} % Punctuation after theorem head % default: no punctuation
  { } % Space after theorem head
  {} % Theorem head spec
\theoremstyle{theorems}
\newtheorem{eodefinition}{Definition}
\newtheorem{eotheorem}{Theorem}

\addbibresource{refs.bib}

\newcommand\br{\\[-4pt]}
\newcounter{rule}
\renewcommand\therule{R\arabic{rule}}
\newcommand\rr{\smalltriangleright{}}
\newcommand\rrule[1]{{\scshape\sffamily\ref{rule:#1}}}
\newcommand{\jrule}[1]{%
  \refstepcounter{rule}\label{rule:#1}%
  \text{\textbf{\rrule{#1}}}}

\newcommand\sgm[1]{$\sigma_#1$}
\newcommand\dlt[1]{$\delta_#1$}
\newcommand\m[1]{$\mu_#1$}

\title{Elimination Of Jump Objects}
\author{author}

@@

\section{Introduction}
The problem of eliminating Jump-like statements is known for both computer science and software engineering communities for a few decades and formulated by several scientists, some of which have suggested solutions for this problem.

The presence of jump statements in the source code makes it impossible to use the functional method of analysis.
Other program analysis methods, including those based on symbolic execution techniques, will also benefit from the removal of jump statements, which are well-known obstacles to the precision of analysis, as explained by \citet{baldoni2018survey}.

There are an "object-flow" languages where the execution flow is controlled by object composition.
An example of such a language is EO, which is based on $\varphi$-calculus introduced by \citet{bugayenko2021eolang} and formalized by \citet{kudasov2021}.
This is a strictly object-oriented programming language in which any action is described by creating a new object, or decorating an existing one.

We are trying to answer the following research question: "Is it possible for any program, written in 'object-flow' language like EO and containing GOTO objects, to find a semantically equivalent program, but without GOTO objects?"

\section{Related work}
The problem was firstly formulated by \citet{dijkstra1968letters}: he affirmed that the use of GOTO has disastrous effect and proposed to remove it from all programming languages.
Even though a few years later \citet{knuth1974structured} showed a manner for structuring programs with the GOTO statement and argued that it could be a powerful tool if it is well used, researchers got interested in finding a way for restructuring programs and eliminating GOTO statements \citep{arsac1977construction}.

To solve the problem of eliminating Jump-like statements in other programming languages, the following works were published:
\Citet{williams1985restructuring} created a tool for restructuring Pascal code;
\Citet{jonsson1989next} was one of the first who made a summary of so called GOTO-patches, which are design techniques for eliminating GOTO statements;
\Citet{erosa1994taming} suggested an algorithm for automatic elimination of GOTO statements in~C;
\Citet{zegour1994new} explained how GOTO elimination may work in any language, as long as it has jump statements and control structures;
\Citet{morris1997goto} suggested to use regular expressions for replacing GOTO with IF-THEN-ELSE constructs;
\Citet{ramshaw1988eliminating} proposed an algorithm for Pascal programs;
\Citet{ganapathi2008shim} introduced optimization method for SHIM IR;
\Citet{ceccato2008goto} suggested how to eliminate GOTO during migration of legacy code to Java.

However, to our knowledge, there is still no method available, which enables automatic elimination of jump objects in programs written in such an object-flow languages.

\section{Definitions and designations}
\begin{eodefinition}
\emph{Dataization} in $\varphi$-calculus is a process of retrieving data from an object.
The dataization of an object at the highest level of composition is similar to the execution of a program in other languages.
\end{eodefinition}
\begin{eodefinition} \emph{Nesting} of the current object determines the number of objects that encapsulate current object via abstraction, application, or decoration directly or through other objects.
For example, here object $x$ has 3-rd level of nesting:
\begin{ffcode}
o1
  o2
  ...
  o3
    o4
      x
\end{ffcode}
\end{eodefinition}
\begin{eodefinition}
\emph{Forwarding goto} denotes an object \ff{goto}, which encapsulate object \ff{forward} directly or through other objects:
\begin{ffcode}
goto
  [o1]
  |$\leadsto$|o1.forward o
\end{ffcode}
\end{eodefinition}
\begin{eodefinition}
\emph{Backwarding goto} denotes an object \ff{goto}, which encapsulate object \ff{backward} directly or through other objects:
\begin{ffcode}
goto
  [o1]
  |$\leadsto$|o1.backward
\end{ffcode}
\end{eodefinition}

Let's look at an EO code example that demonstrates the general syntax of the language:
\begin{ffcode}
memory 0 > x
[] > app
  [a b] > max
    goto > @
      [g]
        if. > @
          a.gt b
          g.forward a
          b
  [x] > run
    while. > @!
      and.
        x.lt 10
        (x.mod 1).eq 0
      [i]
        x.write > @
          (x.plus 1)
\end{ffcode}
In this code snippet, \ff{goto} is an \emph{atom}-object with attribute \ff{g}, which dataizes similar to \ff{seq} object;
\ff{memory} is mutable object, which allows to reinitialize the bound object \ff{x} via \ff{.write} attribute;
\ff{eq} is an object, which checks the equality of the objects passed to it;
\ff{if} is an object, which implements behavior of branching;
\ff{seq} is an object that sequentially dataizes all objects within it;
\ff{while} is an object, implementing behavior of loops: its first bound attribute implements condition of loop dataization and second one implies "body" of loop;
\ff{backward} and \ff{forward} are objects, which implements behaviour of jumps: first of them starts dataization of object \ff{g} again, while the second one returns the object passed to it as a parameter and terminates the dataization of the \ff{goto} object.

In terms of $\varphi$-calculus, code snipped above will be presented as:
\begin{equation}
\begin{split}
& app \mapsto \llbracket \br
& \quad max(a, b) \mapsto \llbracket \br
& \quad \quad \varphi \mapsto \ff{goto}(\llbracket\,g\,\mapsto\,\varnothing,\br
& \quad \quad \quad \varphi \mapsto a.\ff{gt}(b).\ff{if}(g.\ff{forward}(a), b) \br
& \quad \quad \rrbracket) \br
& \quad \rrbracket \br
& \quad run(x) \mapsto \llbracket \br
& \quad \quad \varphi\ff{!} \mapsto x.\ff{lt}(10).\ff{and}(x.\ff{mod}(1).\ff{eq}(0)) \br
& \quad \quad \quad \quad .\ff{while}(\llbracket i \mapsto \varnothing, \br
& \quad \quad \quad \quad \quad \quad \quad \quad \varphi \mapsto x.\ff{write}(x.\ff{plus}(1)) \br
& \quad \quad \quad \quad \quad \rrbracket) \br
& \quad \rrbracket \br
& \rrbracket \\
\end{split}
\end{equation}

\emph{Designations:}
\begin{itemize}
\item "$>$" is used to name objects, where on the left side of ">" is object itself and its name on the right side,
\item "$[]$" is used to specify object parameters,
\item "$@$" is used to denote the object being decorated,
\item "$!$" is used to declare the object as a constant,
\item "$\mapsto$" is used to denote an object, where free attributes stay in the parentheses on the left side and pairs, which represent bound attributes, stay on the right side, in double-square brackets,
\item "$\leadsto$" denotes an arbitrary \emph{nesting} of an object, i.e. it can be encapsulated by other objects directly or via chain of objects.
\end{itemize}

\section{Sample}
Let's consider the following example of C++ code, which uses goto:
\begin{ffcode}
int x = 0;|$\label{ln:cpp0}$|
int func() {
l1:
  x = x + 1;
  if (x < 10)
    goto l1;
  else
    cout << "X >= 10!\n";
  return x;
}|$\label{ln:cpp1}$|
\end{ffcode}

We provide four independent transformations that replace \ff{goto} with a semantically equivalent analogue.
After applying them, example above respresents as:
\begin{ffcode}
int fl_1 = 0;|$\label{ln:tcpp0}$|
int fl_2 = -1;
int x = 0;
int func() {
  fl_1 = 0;
  fl_2 = -1;
  do {
    fl_1 = 1;
    fl_2 = 0;
    x = x + 1;
    if (x < 10)
      fl_1 = 0;
    else
      cout << "X >= 10!\n";
    if (fl_1 == 1)
      fl_2 = 1;
  } while (fl_1 == 0 |$\vert$||$\vert$| fl_2 == -1);
  if (fl_2 == 1)
    return x;
}|$\label{ln:tcpp1}$|
\end{ffcode}

Since we consider the problem of eliminating Jump-like statements on the example of a strictly object-oriented language EO, lines ~\ref{ln:cpp0}--~\ref{ln:cpp1} will be represented as:
\begin{ffcode}
memory 0 > x
[] > func
  goto > @
    [g]
      seq > @
        x.write (x.plus 1)
        if.
          x.lt 10
          g.backward
          QQ.io.stdout "X >= 10!\n"
        g.forward x
\end{ffcode}

While this code after transformations will appear as:
\begin{ffcode}
memory 0 > fl_1
memory -1 > fl_2
memory 0 > x
[] > func
  seq > @
    fl_1.write 0
    fl_2.write -1
    while. > rem!
      or.
        fl_1.eq 0
        fl_2.eq -1
      [i]
        seq > @
          fl_1.write 1
          fl_2.write 0
          x.write (x.plus 1)
          if.
            x.lt 10
            fl_1.write 0
            QQ.io.stdout "X >= 10!\n"
          if.
            fl_1.eq 1
            fl_2.write 1
            TRUE
    if.
      fl_2.eq 1
      x
      rem
\end{ffcode}

Thus, the main purpose of our research is:
\begin{enumerate}
    \item To formulate inference rules for the transformations that we propose,
    \item proof that these 4 transformations are necessary and sufficient to solve the problem under consideration,
    \item suggest a program tool, which
    enables automatic elimination of jump objects in programs written in EO.
\end{enumerate}

\section{Method}
In this article, we will describe an algorithm that will convert an EO program in such a way that the output returns a semantically equivalent program, but without using the GOTO object.

Some objects in EO may need to be platform specific and can’t be composed from other existing objects — they are called \emph{atoms}. 
All jump statements are implemented as a
single \emph{atom}-object \ff{goto}.
The implementation of \ff{goto} in EO implies two types of jumps - forward jump and backward jump.
It is required to cover all possible cases for both of them.

%%The proposed algorithm is a series of transformations that are performed depending on the type of "jump".
In total, we describe 4 types of transformations:
\begin{inparaenum}[1)]
\item \emph{Simple goto} (further $SG$),
\item \emph{Goto to While} (further $GTW$) and
\item \emph{Goto forward/backward} (further $GF$/$GB$).
\end{inparaenum}
Transformations are independent and not sequential, and each one is executed in its own particular case.
For example, $SG$ and $GTW$ are only a "preparation" for the actual replacement of the jump by a semantically equivalent analog.
First one is executed if there is no condition before the jump itself, while second one transforms object \ff{goto} into object \ff{while}.
$GF$/$GB$ are the main transformations, as they contain the jump replacement itself and cover most possible cases of using forward and backward jumps.
%%\emph{TW} is an "addition" for $GF$ and $GB$ transformations, since the behavior of the \ff{while} object in EO is different from all other objects and dataization of jump inside the loop requires separate consideration.

%%Thus, the algorithm itself consists of the following steps: 
%%\begin{inparaenum}[1)]
%%\item Apply \emph{SF}/\emph{SB} transformation if necessary,
%%\item Find all occurrences of forward/backward jump's calls,
%%\item Apply $GF$/$GB$ transformations to all of them,
%%\item If there is a \emph{while.} object that wraps the jump call, apply the \emph{WF}/\emph{WB} transformation.
%%\end{inparaenum}

The combination of all 4 transformations implies a complete algorithm for getting rid of the GOTO object, as they cover all possible cases of using GOTO and forward/backward jumps in the language.

Since we consider a solution to the problem in strictly object-oriented programming languages, which means that the whole program is a set of objects where some objects can encapsulate others.
It should be noted that each object in the program has its own level of \emph{nesting} relative to other objects.

We will also consider our transformations in the form of $\varphi$-calculus, on which EO is based.
The proposed $\varphi$-calculus represents an object model through data and objects, while operations with them are possible through abstraction, application and decoration.
The calculus introduces a formal apparatus for manipulations with objects.

\subsection{Simple goto} \label{SG}
As mentioned above, this transformation is a "preparation" for $GF$/$GB$ which are make the replacement of object \ff{goto} itself.
Since $GF$/$GB$ only consider the use of the \ff{forward}/\ff{backward} objects when they are directly encapsulated in the \ff{if} object(\cref{GF}/\cref{GB}), it is required to add an additional \ff{if} when it is absent before the jump.
Thus, the transformation under consideration adds an additional fictitious condition $\emph{if (true)}$ if it necessary.

Let's consider the case where the \ff{forward} is not directly encapsulated in the \ff{if} object:
\begin{ffcode}
goto
  [g]
   |$\leadsto$|g.forward y|$\label{ln:sf1}$|
\end{ffcode}
Here, $\alpha_{i}$, $x$ and $y$ defines any object in the language, excluding \ff{if}.

Applying the transformation, the example above maps to:
\begin{ffcode}
goto
  [g]
   |$\leadsto$|if.|$\label{ln:tsf0}$|
      TRUE
      g.forward y
      TRUE|$\label{ln:tsf1}$|
\end{ffcode}

Describing transformation in terms of $\varphi$-calculus, code at line ~\ref{ln:sf1} will be represented as:
\begin{equation}
\begin{split}
& g.\ff{forward}( y ) \\
\end{split}
\end{equation}
and code at lines ~\ref{ln:tsf0}---\ref{ln:tsf1} as:
\begin{equation}
\begin{split}
& \ff{TRUE}.\ff{if}(g.\ff{forward}( y )) \\
\end{split}
\end{equation}

\begin{eotheorem}
Any object is semantically equivalent to the same object encapsulated in \ff{TRUE}.\ff{if}:
\begin{equation*}
\forall x (x \equiv \ff{TRUE}.\ff{if}(x)).
\end{equation*}
\end{eotheorem}

\begin{proof}
By definition of the language, if there is any object \emph{Z}, then the following equivalence follows:
\begin{equation}
Z \equiv if(true,\,Z,\,true),
\end{equation}
where first argument in \emph{if} is condition itself, second argument dataizes when condition is \emph{true} and third argument dataizes when condition is \emph{false}.
\end{proof}

\begin{equation*}
\dfrac
    {\begin{split}
    & g.\ff{forward}( y ) \\
    \end{split}}
    {\begin{split}
    & \ff{TRUE}.\ff{if}(g.\ff{forward}( y )) \\
    \end{split}}
    \jrule{1}
\end{equation*}

%%Note that in this example, \ff{forward} is simply an attribute of the abstract object \emph{g} and is dataized as soon as the queue reaches it.
In such cases where there is an "unconditional" jump, first of all we transform it into "conditional" jump by encapsulating it in additional object \ff{if}---\emph{if (true)}.
This will not violate the program logic, since added condition will always be \emph{TRUE} and "THEN" brunch of if-statement with object \ff{forward} will always be dataized.

The same logic applies with the \ff{backward} object.

\subsection{Goto forward} \label{GF}
In this section, we discuss the transformation by replacing \emph{forwarding goto} with a semantic equivalent.
Based on the concept of transformation, only cases in which \ff{forward} is directly encapsulated in the object \ff{if} are considered.

The essence of the transformation is that we create an additional $flag$ object which will assign as "0" or "1" and will show whether the jump condition has been met.
Instead of dataization \ff{forward}, we just reinitialize the $flag$ object with another value of data object \ff{Int}.
Moreover, it is necessary to protect from dataization all objects inside \ff{goto} that follows the jump, since they should not be dataized when the jump condition is \ff{TRUE}.
This requires encapsulating each such object with an additional object \ff{if} to check the equality $flag$ to an object "1"---i.e. whether the jump condition has been met before or not.

Let's consider the following example:
\begin{ffcode}
goto|$\label{ln:of0}$|
  [g]
   |$\leadsto$|if.
      k
      g.forward y
      |$\alpha_{n}$||$\label{ln:of1}$|
\end{ffcode}
Here, $\alpha_i$, $x$ and $y$ defines any object in the language; $k$ defines an object, which when dataized equals to a data object \ff{TRUE}/\ff{FALSE}.

Applying the transformation, the example above maps
to:
\begin{ffcode}
memory -1 > flag
seq|$\label{ln:tof0}$|
  goto > rem!
    [g]
     |$\leadsto$|if.
        k
        flag.write 1
        |$\alpha_{n+1}$|
  if.
    flag.eq 1
    y
    rem|$\label{ln:tof1}$|
\end{ffcode}

In terms of $\varphi$-calculus lines ~\ref{ln:of0}--\ref{ln:of1} appears as:
\begin{equation}
\begin{split}
& \ff{goto}(\llbracket g\,\mapsto\,\varnothing,\br
& \quad \quad \quad \varphi\,\mapsto\,(\,\leadsto k.\ff{if}(g.\ff{forward}(y), \alpha_{n+1}), \br
& \quad \quad \quad \rrbracket) \\
\end{split}
\end{equation}
Lines ~\ref{ln:tof0}--\ref{ln:tof1} will be represented as:
\begin{equation}
\begin{split}
& \ff{seq}(\ff{goto}(\llbracket g\,\mapsto\,\varnothing,\br
& \quad \quad \quad \quad \quad \varphi\,\mapsto\,(\,\leadsto x(\alpha_1\,\ldots\,\alpha_n, \br
& \quad \quad \quad \quad \quad \quad \quad k.\ff{if}(flag.\ff{write}(1),\,\alpha_{n+1}), \br
& \quad \quad \quad \quad \quad \quad \quad \Theta(\alpha_{n+2}),\,\ldots\,,\,\Theta(\alpha_m)),\ldots\Theta(\alpha_{m+1})) \br
& \quad \quad \quad \quad \quad \rrbracket)\mapsto rem\ff{!}, \br
& \quad \quad flag.\ff{eq}(1).\ff{if}(y,\,rem)) \\
\end{split}
\end{equation}
Here,
\begin{equation}
\Theta(X) \equiv flag.\ff{eq}(0).\ff{if}(X,\,\ff{TRUE})
\end{equation}

\begin{eotheorem}
\emph{Forwarding goto} maps to semantically equivalent analog without \ff{forward} objects:
\begin{equation*}
\begin{split}
& \ff{goto}(\llbracket g\,\mapsto\,\varnothing,\,\varphi\,\leadsto\,x(\ldots,\,\leadsto g.\ff{forward}(y),\,\ldots)\rrbracket) \\
& \downarrow \\
& \ff{seq}( \br
& \quad \quad \ff{goto}(\llbracket g\,\mapsto\,\varnothing,\br
& \quad \quad \quad \quad \quad \varphi\,\leadsto\,x(\ldots,\,\leadsto flag.\ff{write}(1),\,\ldots)\br
& \quad \quad \quad \quad \quad \rrbracket)\mapsto rem\ff{!}, \br
& \quad \quad flag.\ff{eq}(1).\ff{if}(y,\,rem)) \\
\end{split}
\end{equation*}
\end{eotheorem}

\begin{proof}
Let's introduce the following recursive function:
\begin{equation}
\begin{split}
(\forall x\,\in\,\emph{attr}(\emph{g})):\,\mu(x)\,&\rightarrow\,(\Theta(x),\,\mu(x'\,\neq\,x)) \\
        &\rightarrow\,(\Theta(x),\,(\Theta(x'),\,\mu(x''))) \\
        &\ldots
\end{split}
\end{equation}
Here, $attr(X)$ denotes all attributes of an object $X$.

Since the object \ff{goto} without \ff{forward} inside dataizes similar to the object \ff{seq}, then it follows that the following expression is valid:
\begin{equation}
\begin{split}
& \ff{goto}(\llbracket g\,\mapsto\,\varnothing,\br
& \quad \quad \varphi\,\leadsto\,x(\alpha_1,\,\ldots\,,\alpha_n)\rrbracket) \equiv \\
& \ff{seq}(\ff{goto}(\llbracket g\,\mapsto\,\varnothing,\br
& \quad \quad \varphi\,\leadsto\,x(\alpha_1,\,\ldots\,,\alpha_n)\rrbracket)),\\
\end{split}
\end{equation}
Here, $alpha_n$ defines any object in the language, excluding object \ff{forward}.

When the outer \ff{seq} dataizes object \ff{goto} and the jump condition is \ff{TRUE}, it must return as a result of dataization the attribute of an object \ff{forward}.
It can be done by the \ff{if} object, that follows object \ff{goto} and checks equality of $flag$ object to "1".

According to the logic of dataization and definition of the language, both cases of dataization object \ff{forward} are considered via added \ff{if} objects.
\debate[Mikhail]{Need a STRONG proof!}
\end{proof}

\begin{equation*}
\dfrac
    {\begin{split}
    & \ff{goto}(\llbracket\,g\,\mapsto\,\varnothing,\br
    & \quad \quad \quad \varphi\,\leadsto\,x(\ldots,\,k.\ff{if}(g.\ff{forward}(y),\,z),\br
    & \quad \quad \quad w,\,\ldots)\rrbracket) \\
    \end{split}}
    {\begin{split}
    & \ff{seq}(\ff{goto}(\llbracket g\,\mapsto\,\varnothing,\br
    & \quad \quad \quad \quad \quad \varphi\,\leadsto\,x(\ldots,\,k.\ff{if}(flag.\ff{write}(1),\,z), \br
    & \quad \quad \quad \quad \quad \quad \quad \mu(w),\,\ldots\,) \br
    & \quad \quad \quad \quad \quad \rrbracket)\mapsto rem\ff{!}, \br
    & \quad \quad flag.\ff{eq}(1).\ff{if}(y,\,rem)) \\
    \end{split}}
    \jrule{2}
\end{equation*}
Here, $x$, $w$, $y$ and $z$ defines any object in the language, $k$ defines an object, which when dataized equals to a data object \ff{TRUE}.

Indeed, depending on the jump condition, the same objects will be dataized as when dataization the \ff{forward} object itself.

It is not difficult to notice that objects which goto-forward "jumps through" will be dataized if the condition in the if-statement is not equal to \ff{TRUE}.
So before dataization such each object, that follows the jump, we can check condition of the jump.
It can be done via mutable object \ff{memory} and additional \ff{if} objects.
This is required so that all objects following the jump that are attributes of the abstract object $g$ straightforwardly or throw other objects are not dataized.
So, by encapsulating each object, following the jump and whose \emph{nesting} is at least that of the $g$ object, in this additional \ff{if} object with a $flag$ check, we make sure that these objects will not be dataized when the jump condition is met.

\subsection{Goto to While} \label{GTW}
As mentioned above, all transformations are sequential.
A necessary condition for the $GTW$ transformation is the absence of \ff{forward}/\ff{backward} objects inside the \ff{goto} object.
In other words, this transformation is applied directly along with $GF$ and $GB$, since they imply the replacement of \ff{forward} and \ff{backward} objects.
The essence of the transformation is that the object \ff{goto} maps to a control-flow object \ff{while}.

Let's consider an example where inside the \ff{goto} object, the \ff{forward}/\ff{backward} objects are replaced with a semantic equivalent:
\begin{ffcode}
memory -1 > flag
seq|$\label{ln:gtw0}$|
  goto > rem!
    [g]
     |$\leadsto$|x
        |$\alpha_1$|
        ...
        |$\alpha_n$|
        if.
          k
          flag.write 1
          |$\alpha_{n+1}$|
        if.
          flag.eq 0
          |$\alpha_{n+2}$|
          TRUE
        ...
        if.
          flag.eq 0
          |$\alpha_{m}$|
          TRUE
  if.
    flag.eq 1
    y
    rem|$\label{ln:gtw1}$|
\end{ffcode}
Here, $\alpha_{i}$, $y$ and $x$ defines any object in the language, excluding \ff{forward}/\ff{backward};
$k$ defines an object, which when dataized equals to a data object \ff{TRUE}/\ff{FALSE}.

Applying the transformation, the example above maps
to:
\begin{ffcode}
memory -1 > flag
seq|$\label{ln:tgtw0}$|
  while. > rem!
    flag.eq -1
    [i]
      seq > @
        flag.write 0
        |$\leadsto$|x
          |$\alpha_1$|
          ...
          |$\alpha_n$|
          if.
            k
            flag.write 1
            |$\alpha_{n+1}$|
          if.
            flag.eq 0
            |$\alpha_{n+2}$|
            TRUE
          ...
          if.
            flag.eq 0
            |$\alpha_{m}$|
            TRUE
  if.
    flag.eq 1
    y
    rem|$\label{ln:tgtw1}$|
\end{ffcode}

In terms of $\varphi$-calculus lines ~\ref{ln:gtw0}--\ref{ln:gtw1} appears as:
\begin{equation}
\begin{split}
& \ff{seq}(\ff{goto}(\llbracket g\,\mapsto\,\varnothing,\br
& \quad \quad \quad \quad \quad \varphi\,\mapsto\,(\,\leadsto x(\alpha_1\,\ldots\,\alpha_n, \br
& \quad \quad \quad \quad \quad \quad \quad k.\ff{if}(flag.\ff{write}(1),\,\alpha_{n+1}), \br
& \quad \quad \quad \quad \quad \quad \quad \Theta(\alpha_{n+2}),\,\ldots\,,\,\Theta(\alpha_m))) \br
& \quad \quad \quad \quad \quad \rrbracket)\mapsto rem\ff{!}, \br
& \quad \quad flag.\ff{eq}(1).\ff{if}(y,\,rem)) \\
\end{split}
\end{equation}
Here, $alpha_n$ defines any object in the language, excluding objects \ff{forward} and \ff{backward}.
